---
title: "Quantium Virtual Internship - Retail Strategy and Analytics - Task 1"
author: "Emmanuel Aina"
date: "2025-06-02"
output: 
  pdf_document:
    df_print: default
    highlight: tango
    keep_tex: yes
    latex_engine: pdflatex
mainfont: "Roboto"
monofont: "Consolas"
header-includes:
  - \usepackage{fvextra}
  - \usepackage{graphicx}
  - \usepackage{setspace}
  - \setstretch{1.5}
---

```{r setup, include=FALSE}
# Set options for R Markdown knitting
knitr::opts_chunk$set(echo = TRUE, linewidth = 80)

# Load required libraries
library(knitr)

# Set up line wrapping in output
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (!is.null(options$linewidth)) {
    x <- knitr:::split_lines(x)
    if (any(nchar(x) > options$linewidth)) x <- strwrap(x, width = options$linewidth)
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
```
# This file is a solution template for the Task 1 of the Quantium Virtual Internship.

## Load required libraries and datasets
Note that you will need to install these libraries if you have never used these before.

```{r 0 Load Libraries, results ='hide'}
library(data.table)
library(ggplot2)
library(ggmosaic)
library(readr)
purchase_behaviour <- read_csv("QVI_purchase_behaviour.csv")
transaction_data <- read_csv("QVI_transaction_dataa.csv")
```
## Exploratory data analysis
The first step in any analysis is to first understand the data. Let's take a look at each of the data sets provided.

### Examining transaction_data
```{r Examining transaction data}
#### Load necessary libraries
library(data.table)

#### Check the structure of the data set
str(transaction_data)

#### Display the first 10 rows
head(transaction_data, 10)

#### Check if numeric columns are indeed numeric
sapply(transaction_data, class)

#### Display summary statistics
summary(transaction_data)
```
#### Let's change this to a date format.
```{r Convert DATE to date format}
#### Convert DATE column to a date format
transaction_data$DATE <- as.Date(transaction_data$DATE, format = "%d/%m/%y")
```

#### We should check that we are looking at the right products by examining PROD_NAME
```{r Summary of PROD_NAME}
#### Check the structure of the column
str(transaction_data$PROD_NAME)

#### View the first few product names
head(transaction_data$PROD_NAME, 10)

#### Count unique product names
num_unique_products <- length(unique(transaction_data$PROD_NAME))
num_unique_products

#### Display summary statistics
summary(transaction_data$PROD_NAME)
```
Looks like we are definitely looking at potato chips but how can we check that these are all chips?
We can do some basic text analysis by summarizing the individual words in the product name.

#### Further examine on PROD_NAME
```{r Further examine PROD_NAME}
#### Ensure correct data set reference (adjust as needed)
productWords <- data.table(unlist(strsplit(unique(transaction_data$PROD_NAME), " ")))

#### Rename the column to 'words'
setnames(productWords, "words")

#### Count frequency of words
wordCounts <- productWords[, .N, by = words][order(-N)]

#### View the most common words
head(wordCounts, 20)
```
As we are only interested in words that will tell us if the product is chips or not, let's remove all words with digits and special characters such as '&' from our set of product words.

#### Remove digits, and special characters, and then sort the distinct words by frequency of occurrence.
```{r Clean Product Words}
# Extract words from product names
productWords <- data.table(unlist(strsplit(unique(transaction_data$PROD_NAME), " ")))
setnames(productWords, "words")

# Remove words containing digits (0-9) or special characters ('&', '@', etc.)
cleanWords <- productWords[!grepl("[0-9&@#$%^*()_+=\\-]", words), ]

# Count frequency of cleaned words
wordCounts <- cleanWords[, .N, by = words][order(-N)]

# Display the most common words after cleaning
head(wordCounts, 20)
```
There are salsa products in the data set but we are only interested in the chips category, so let's remove these.

#### Remove salsa products
```{r}
# Create a logical vector to identify products containing "salsa"
transaction_data$SALSA <- grepl("salsa", tolower(transaction_data$PROD_NAME))

# Remove rows where SALSA is TRUE (i.e., products containing "salsa")
transaction_data <- transaction_data[transaction_data$SALSA == FALSE, ]

# Drop the SALSA column since it's no longer needed
transaction_data$SALSA <- NULL

# Display the first few rows of the cleaned data set
head(transaction_data)
```
There are no nulls in the columns, so lets check for possible outliers

#### Summaries the data to check for possible outliers
```{r initial summary}


# Generate summary statistics for all columns
summary(transaction_data)

# Check for null values explicitly in each column
colSums(is.na(transaction_data))
```
There are no nulls in the columns but product quantity appears to have an outlier which we should investigate further. Let's investigate further the case where 200 packets of chips are bought in one transaction.

```{r Investigate Customer Transactions}
# Identify the customer who purchased 200 packets of chips
customer_id <- transaction_data$LYLTY_CARD_NBR[transaction_data$PROD_QTY == 200]

# Filter transactions made by this customer
customer_transactions <- transaction_data[transaction_data$LYLTY_CARD_NBR == customer_id, ]

# View a summary of their transactions
summary(customer_transactions)

# Display sample transactions
head(customer_transactions)
```
There are two transactions where 200 packets of chips are bought in one transaction and both of these transactions were by the same customer.
It looks like this customer has only had the two transactions over the year and is  not an ordinary retail customer. The customer might be buying chips for commercial purposes instead. We'll remove this loyalty card number from further analysis.

#### Filter out the customer based on the loyalty card number
```{r Filter Customer Transactions}
# Identify the customer with bulk purchases
customer_id <- transaction_data$LYLTY_CARD_NBR[transaction_data$PROD_QTY == 200]

# Remove all transactions from this customer
filtered_transaction_data <- transaction_data[transaction_data$LYLTY_CARD_NBR != customer_id, ]

# Re-examine the dataset after filtering
summary(filtered_transaction_data)

# Display the first few rows
head(filtered_transaction_data)
```
That's better. Now, let's look at the number of transaction lines over time to see if there are any obvious data issues such as missing data.

#### Count Transaction by date
```{r Transaction Count by Date}
# Count transactions for each unique date
transaction_summary <- aggregate(TXN_ID ~ DATE, data = transaction_data, FUN = length)

# Rename columns for clarity
colnames(transaction_summary) <- c("Date", "Transaction_Count")

# View summary of transactions per date
summary(transaction_summary)

# Display the first few rows
head(transaction_summary)
```
There's only 364 rows, meaning only 364 dates which indicates a missing date. Let's create a sequence of dates from 1 Jul 2018 to 30 Jun 2019 and use this to create a chart of number of transactions over time to find the missing date.

#### Created a sequence of dates and join this the count of transactions by date
```{r, echo=FALSE, fig.align='center'}
library(ggplot2)

# Step 1: Create a complete sequence of dates
full_dates <- data.frame(DATE = seq(as.Date("2018-07-01"), as.Date("2019-06-30"), by = "day"))

# Step 2: Count transactions per available date
transactions_by_day <- aggregate(TXN_ID ~ DATE, data = transaction_data, FUN = length)
colnames(transactions_by_day) <- c("DATE", "Transaction_Count")

# Step 3: Merge full date sequence with transaction counts
final_summary <- merge(full_dates, transactions_by_day, by = "DATE", all.x = TRUE)

# Step 4: Identify missing date
missing_dates <- final_summary$DATE[is.na(final_summary$Transaction_Count)]
print(missing_dates)  # This will output the missing date

# Step 5: Fill missing transaction counts with 0
final_summary$Transaction_Count[is.na(final_summary$Transaction_Count)] <- 0

# Step 6: Plot transactions over time
ggplot(final_summary, aes(x = DATE, y = Transaction_Count)) +
  geom_line() +
  labs(x = "Day", y = "Number of transactions", title = "Transactions Over Time") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

head(transactions_by_day)
summary(transactions_by_day)
```
We can see that there is an increase in purchases in December and a break in Late December, Let's zoom in on this.

#### Filtered to December and look at individual days
```{r, echo=FALSE, fig.align='center'}
library(ggplot2)

# Filter the final_summary data for December 2018
december_data <- final_summary[final_summary$DATE >= as.Date("2018-12-01") & 
                               final_summary$DATE <= as.Date("2018-12-31"), ]

# Alternatively using dplyr (if you prefer):
# library(dplyr)
# December_data <- final_summary %>% filter(DATE >= as.Date("2018-12-01") & DATE <= as.Date("2018-12-31"))

# Plot transactions over time for December 2018
ggplot(december_data, aes(x = DATE, y = Transaction_Count)) +
  geom_line() +
  labs(x = "Day", y = "Number of Transactions", title = "Transactions Over Time") +
  scale_x_date(date_breaks = "1 day", date_labels = "%d %b") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
We can see that the increase in sales occurs in the lead-up to Christmas and that there are zero sales on Christmas day itself. This is due to shops being closed on Christmas day.

#### created a new column 'PACK_SIZE' by extracting the first numeric value from PROD_NAME
```{r}
transaction_data$PACK_SIZE <- parse_number(transaction_data$PROD_NAME)

# Create a frequency table of PACK_SIZE
pack_summary <- as.data.frame(table(transaction_data$PACK_SIZE))

# Rename the columns for readability
names(pack_summary) <- c("PACK_SIZE", "Frequency")

# Convert PACK_SIZE from factor to numeric, if necessary
pack_summary$PACK_SIZE <- as.numeric(as.character(pack_summary$PACK_SIZE))

# Order the summary by pack size
pack_summary <- pack_summary[order(pack_summary$PACK_SIZE), ]

# Print the summary to check if the pack sizes look sensible
print(pack_summary)
```
The largest size is 380g and the smallest size is 70g - seems sensible!


#### Created plot a histogram of PACK_SIZE since we know that it is a categorical variable and not a continuous variable even though it is numeric.
```{r, echo=FALSE, fig.align='center'}
# If PACK_SIZE isn't already a factor, convert it so that ggplot treats it  as categorical:
 transaction_data$PACK_SIZE<-factor(transaction_data$PACK_SIZE)

# Load ggplot2
library(ggplot2)

# Create a bar chart of PACK_SIZE:
library(ggplot2)

# Create a bar chart of PACK_SIZE:
ggplot(transaction_data, aes(x = PACK_SIZE)) +
  geom_bar(fill = "skyblue", color = "black", stat = "count") +
  labs(x = "Pack Size", y = "Number of Transactions", title = "Transactions by Pack Size") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center the title
```
Pack sizes created look reasonable. 

#### Created brands, we can use the first word in PROD_NAME to work out the brand name
```{r}
# Create the BRAND column by extracting the first word from PROD_NAME
transaction_data$BRAND <- sub(" .*", "", transaction_data$PROD_NAME)

# Check the results by creating a frequency table of the brands
brand_summary <- as.data.frame(table(transaction_data$BRAND))
names(brand_summary) <- c("BRAND", "Frequency")

# Order by frequency (descending) to see the most common brands first
brand_summary <- brand_summary[order(-brand_summary$Frequency), ]

# Print the brand summary
print(brand_summary)
```
Some of the brand names look like they are of the same brands - such as RED and RRD, which are both Red Rock Deli chips. Let's combine these together.

#### Cleaned and examined BRAND names 
```{r}
# Clean brand names: for example, change "RED" to "RRD"
transaction_data$BRAND[transaction_data$BRAND == "Red"] <- "RRD"
transaction_data$BRAND[transaction_data$BRAND =="SNBTS"] <-"SUNBITES"
transaction_data$BRAND[transaction_data$BRAND =="INFZNS"]<- "INFUZIONS"
transaction_data$BRAND[transaction_data$BRAND =="WW"]<-"WOOLWORTHS"
transaction_data$BRAND[transaction_data$BRAND =="SMITH"]<- "SMITHS"
transaction_data$BRAND[transaction_data$BRAND =="NCC"]<- "NATURAL"
transaction_data$BRAND[transaction_data$BRAND =="DORITO"]<- "DORITOS"
transaction_data$BRAND[transaction_data$BRAND =="GRAIN"]<- "GRNWVES"
 
# Check the cleaned results by creating a frequency table of the brands
brand_summary <- as.data.frame(table(transaction_data$BRAND))
names(brand_summary) <- c("BRAND", "Frequency")

# Order the summary by frequency (optional: descending order)
brand_summary <- brand_summary[order(-brand_summary$Frequency), ]

# Print the summary to see if the brand names now look reasonable
print(brand_summary)
```
Now that i am satisfied with the transaction data set, let's have a look at the customer data set. 

### Examining Purchase_behaviour data
```{r, echo=FALSE, fig.align='center'}

# Examining customer data: purchase_behavior

# Examine the structure and basic summary of the data set
str(purchase_behaviour)
summary(purchase_behaviour)

# Frequency distribution for LIFESTAGE
if("LIFESTAGE" %in% names(purchase_behaviour)) {
  cat("Frequency distribution of LIFESTAGE:\n")
  print(table(purchase_behaviour$LIFESTAGE))
}

# Frequency distribution for PREMIUM_CUSTOMER
if("PREMIUM_CUSTOMER" %in% names(purchase_behaviour)) {
  cat("\nFrequency distribution of PREMIUM_CUSTOMER:\n")
  print(table(purchase_behaviour$PREMIUM_CUSTOMER))
}

# Check the number of unique LYLTY_CARD_NBR values (customer identifiers)
if("LYLTY_CARD_NBR" %in% names(purchase_behaviour)) {
  cat("\nNumber of unique loyalty card numbers:\n")
  print(length(unique(purchase_behaviour$LYLTY_CARD_NBR)))
}

# Visualize the distribution of LIFESTAGE using a bar chart
if("LIFESTAGE" %in% names(purchase_behaviour)) {
  ggplot(purchase_behaviour, aes(x = factor(LIFESTAGE))) +
    geom_bar(fill = "dodgerblue", color = "black") +
    labs(title = "Distribution of LIFESTAGE",
         x = "LIFESTAGE",
         y = "Count") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Visualize the distribution of PREMIUM_CUSTOMER using a bar chart
if("PREMIUM_CUSTOMER" %in% names(purchase_behaviour)) {
  ggplot(purchase_behaviour, aes(x = factor(PREMIUM_CUSTOMER))) +
    geom_bar(fill = "red", color = "yellow") +
    labs(title = "Distribution of Premium Customer Status",
         x = "PREMIUM_CUSTOMER",
         y = "Count") +
    theme_minimal()
}
```
Satisfied with the purchase_behaviour data set

### Merge transaction_data to purchase_behaviour
```{r }
 #### Merge transaction data to customer data
data <- merge(transaction_data, purchase_behaviour, by = "LYLTY_CARD_NBR", all.x = TRUE)
```
```{r Check-for-missing-customer-details, message=FALSE, warning=FALSE}
# Check for missing customer details in merged dataset
missing_customers <- data[is.na(data$LIFESTAGE) | is.na(data$PREMIUM_CUSTOMER), ]

# Print summary of missing records
cat("Number of transactions without a matched customer:", nrow(missing_customers), "\n")

# Optionally, view the first few rows of missing records
if(nrow(missing_customers) > 0) {
  print(head(missing_customers))
}
```
Great, there are no nulls! So all our customers in the transaction data has been accounted for in the customer data set

### Saved file for further analysis 
```{r Code to save dataset as a csv}
fwrite(data, "C:/Users/user/Desktop/QVI_data.csv")

file.exists("C:/Users/user/Desktop/QVI_data.csv")
```
Data exploration is now complete!

Let's start with calculating total sales by LIFESTAGE and PREMIUM_CUSTOMER and plotting the split by these segments to describe which customer segment contribute most to chip sales.

## Total sales by LIFESTAGE and PREMIUM_CUSTOMER
```{r, echo=FALSE, fig.align='center'}
# Load necessary package
library(dplyr)
library(ggplot2)

#Calculate total sales by LIFESTAGE and PREMIUM_CUSTOMER
 total_sales<- data%>%
 group_by(LIFESTAGE,PREMIUM_CUSTOMER)%>%
 summarise(Total_Sales=sum(TOT_SALES))%>%
 arrange(desc(Total_Sales))
 27
# Print summary table
 print(total_sales)

# Create a mosaic plot
p <- ggplot(data = total_sales) +
  geom_mosaic(aes(weight = Total_Sales, 
                  x = product(PREMIUM_CUSTOMER, LIFESTAGE), 
                  fill = PREMIUM_CUSTOMER)) +
  labs(title = "Proportion of the Sale",
       x = "Lifestage",
       y = "Premium Customer Flag") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

#### Improved Label Placement with Proportion of Sales
p + geom_text(data= ggplot_build(p)$data[[1]], aes(x= (xmin + xmax)/2+ 0.01 ,y=
 (ymin + ymax)/2,label= as.character(paste(round(.wt/sum(.wt),3)*100,
 '%'))))
```
Sales are coming mainly from Budget - older families, Mainstream - young singles/couples, and Mainstream - retirees

## Lets's see if the higher sales are due to there being more customers who buy chips.
```{r, echo=FALSE, fig.align='center'}
library(dplyr)
library(ggplot2)
library(ggmosaic)

# Calculate number of unique customers by LIFESTAGE and PREMIUM_CUSTOMER
customer_counts <- data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(Number_of_Customers = n_distinct(LYLTY_CARD_NBR)) %>%
  arrange(desc(Number_of_Customers))

# Print summary table
print(customer_counts)


P <- ggplot(data=customer_counts) +
         geom_mosaic(aes(weight= Number_of_customers, X = product (PREMIUM_CUSTOMER, LIFESTAGE), fill = PREMIUM_CUSTOMER)) +
         labs(title = "Proportion of the Customer",
              x = "Lifestage",
              y = "premium customer flag") +
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5))


p + geom_text(data= ggplot_build(p)$data[[1]], aes(x= (xmin + xmax)/2 + 0.01,y=
 (ymin + ymax)/2,label= as.character(paste(round(.wt/sum(.wt),3)*100,
 '%'))))
```
There are more Mainstream - young singles/couples and Mainstream - retirees who buy chips. This contributes to there being more sales to these customer segments but this is not a major driver for the Budget - Older families segment.
Higher sales may also be driven by more units of chips being bought per customer.

## Average number of units per customer by LIFESTAGE and PREMUIM_CUSTOMER 
```{r, fig.width = 10, fig.align='center'}
# Your ggplot code here}
library(dplyr)
library(ggplot2)


# Calculate the average number of units per customer by LIFESTAGE & PREMIUM_CUSTOMER
avg_units_per_customer <- data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(Average_Units = sum(PROD_QTY) / n_distinct(LYLTY_CARD_NBR)) %>%
  arrange(desc(Average_Units))

# Print summary table
print(avg_units_per_customer)

# Create a bar plot to visualize average units per customer by segment
ggplot(data = avg_units_per_customer, aes(x = LIFESTAGE, y = Average_Units, fill = PREMIUM_CUSTOMER)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Average Units Per Customer by Lifestage & Premium Status",
       x = "Lifestage",
       y = "Average Units Purchased",
       fill = "Premium Customer") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Older families and young families in general buy more chips per customer

## Average price per unit sold  by LIFESTAGE and PREMIUM_CUSTOMER
```{r, fig.width = 10, fig.align='center'}
# Calculate the average price per unit sold by LIFESTAGE & PREMIUM_CUSTOMER
avg_price_per_unit <- data %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(Average_Price = sum(TOT_SALES) / sum(PROD_QTY)) %>%
  arrange(desc(Average_Price))

# Print summary table
print(avg_price_per_unit)

# Create a bar plot to visualize the average price per unit by customer segment
ggplot(avg_price_per_unit, aes(x = LIFESTAGE, y = Average_Price, fill = PREMIUM_CUSTOMER)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Average unit Per price by Lifestage & Premium Status",
       x = "Lifestage",
       y = "Average Price Per Unit",
       fill = "Premium Customer") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Mainstream midage and young singles and couples are more willing to pay more per packet of chips compared to their budget and premium counterparts. This may be due to premium shoppers being more likely to buy healthy snacks and when they buy chips, this is mainly for entertainment purposes rather than their own consumption.This is also supported by there being fewer premium midage and young singles and 
couples buying chips compared to their mainstream counterparts.
As the difference in average price per unit isn't large, we can check if this difference is statistically different.

##  Perform an independent t-test between mainstream vs premium and budget midage and young singles and couples
```{r T-Test-Mainstream-Premium, message=FALSE, warning=FALSE}
library(dplyr)

 #### Perform an independent t‐test between mainstream vs premium and budgetmidage andyoung singles and couples

 # Calculate price per unit
library(dplyr)

library(data.table)
# Ensure QVI_data is a data.table
setDT(data)

# Calculate price per unit
price_per_unit <- data[, price := TOT_SALES / PROD_QTY]

# Perform t-test comparing Mainstream vs Premium/Budget in specified life stages
t.test(
  data[LIFESTAGE %in% c("YOUNG SINGLES/COUPLES", "MIDAGE SINGLES/COUPLES") & PREMIUM_CUSTOMER == "Mainstream", price],
  data[LIFESTAGE %in% c("YOUNG SINGLES/COUPLES", "MIDAGE SINGLES/COUPLES") & PREMIUM_CUSTOMER != "Mainstream", price],
  alternative = "greater"

)
```
The t-test results in a p-value < 2.2e-16, i.e. the unit price for mainstream, young and mid-age singles and couples are significantly higher than that of budget or premium, young and midage singles and couples.


## Deep dive into specific customer segments for insights

We have found quite a few interesting insights that we can dive deeper into.We might want to target customer segments that contribute the most to sales to retain them or further increase sales. Let’s look at Mainstream- young singles/couples. For instance, let’s find out if they tend to buy a particular brand of chips.

```{r}
##Step 1: Prepare Data for Analysis, we need to filter data to Mainstream - Young Singles/Couples and extract brand preferences.

library(dplyr)

 #### Deep dive into Mainstream, young singles/couples
library(data.table)

# Segment Mainstream Young Singles/Couples
segment1 <- data[LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER == "Mainstream"]

# Segment Other Customers
other <- data[!(LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER == "Mainstream")]

# Calculate total quantity per segment
quantity_segment1 <- sum(segment1$PROD_QTY)
quantity_other <- sum(other$PROD_QTY)

# Brand affinity calculations
quantity_segment1_by_brand <- segment1[, .(targetSegment = sum(PROD_QTY) / quantity_segment1), by = BRAND]
quantity_other_by_brand <- other[, .(other = sum(PROD_QTY) / quantity_other), by = BRAND]

# Merge data to compare brand affinity
brand_proportions <- merge(quantity_segment1_by_brand, quantity_other_by_brand, by = "BRAND")

# Calculate affinity ratio
brand_proportions[, affinityToBrand := targetSegment / other]

# Sort brands by affinity
brand_proportions[order(-affinityToBrand)]
```
 We can see that :
 • Mainstream young singles/couples are 23% more likely to purchase Tyrrells chips compared to the rest of the population
 • Mainstream young singles/couples are 56% less likely to purchase Burger Rings compared to the rest of the population
 
Let’s also find out if our target segment tends to buy larger packs of chips.

```{r}
 #### Preferred pack size compared to the rest of the population
library(data.table)

# Ensure 'segment1' and 'other' are data.tables
setDT(segment1)
setDT(other)

# Calculate pack proportions
quantity_segment1_by_pack <- segment1[, .(targetSegment = sum(PROD_QTY) / quantity_segment1), by = PACK_SIZE]
quantity_other_by_pack <- other[, .(other = sum(PROD_QTY) / quantity_other), by = PACK_SIZE]

# Merge data and compute affinity score
pack_proportions <- merge(quantity_segment1_by_pack, quantity_other_by_pack, by = "PACK_SIZE")

# Calculate affinity to pack size
pack_proportions[, affinityToPack := targetSegment / other]

# Sort results by affinity descending
pack_proportions[order(-affinityToPack)]
```

It looks like Mainstream young singles/couples are 27% more likely to purchase a 270g pack of chips compared to the rest of the population but let’s dive into what brands sell this pack size.
```{r}
data[PACK_SIZE == 270, unique(PROD_NAME)]
```
Twisties are the only brand offering 270g packs and so this may instead be reflecting a higher likelihood of purchasing Twisties.

# Conclusion

Let’s recap what we’ve found!

Sales have mainly been due to Budget-older families,Mainstream-youngsingles/couples,and Mainstream- retirees shoppers. We found that the high spend in chips for mainstream young singles/couples and retirees is due to there being more of them than other buyers. Mainstream, midage and young singles and couples are also more likely to pay more per packet of chips. This is indicative of impulse buying behaviour.
We’ve also found that Mainstream youngsingles and couples are 23%more likely to purchase Tyrrells chips compared to the rest of the population. The Category Manager may want to increase the category’s performance by off-locating some Tyrrells and smaller packs of chips in discretionary space near segments where young singles and couples frequent more often to increase visibility and impulse behaviour.

Quantium can help the Category Manager with recommendations of where these segments are and further help them with measuring the impact of the changed placement. We’ll work on measuring the impact of trials in the next task and putting all these together in the third task